### 2024-02-29 doc/manuscript/SI_tables.Rmd
finalize SI tables

---
title: ''
output: 
    #pdf_document
    bookdown::word_document2: 
      reference_docx: word-styles-reference.docx #make sure this line is indented! see this guide for formatting: https://rmarkdown.rstudio.com/articles_docx.html
bibliography: references.bib
csl: new-phytologist.csl

---

```{r eval = TRUE, echo=FALSE, warning=FALSE, results='hide'}

#Appendices
appendix_ForC_updates = "Methods S1"
appendix_IPCC_definitions = "Notes S1"
table_ForCchanges = "Table S1"
table_ForCfieldmapping = "Table S2"
table_variables ="Table S3"

```

**`r table_ForCchanges`. Updates to ForC fields implemented between releases of v3.0 and v4.0.** 
```{r table_ForCchanges, eval = TRUE, echo=FALSE, warning=FALSE, message = FALSE}
library (readr)
urlfile="https://raw.githubusercontent.com/forc-db/ForC/master/database_management_records/record_of_changes.csv"
record_of_changes <-read_csv(url(urlfile))
names(record_of_changes) <- stringr::str_replace_all(names(record_of_changes), "\\.", " ")
record_of_changes <- record_of_changes[c(1:5)] #include all fields but date of change
record_of_changes <- record_of_changes[!grepl("housekeeping", record_of_changes$Motivation, ignore.case = TRUE), c(1:5)]
record_of_changes[is.na(record_of_changes)]<- "" #replace NAs with blanks

library(flextable)
set_flextable_defaults(font.family = "Arial", font.size = 8) # set font
ft <- flextable(record_of_changes)
ft <- autofit(ft)
ft <- width(ft, 1, width = 1)
ft <- width(ft, 2:5, width = 1.5)
ft

```

\clearpage

**`r table_ForCfieldmapping`. Mapping of ForC fields to EFDB.** Details documented in the public GitHub repository associated with the project, IPCC-EFDB-integration repository within the ForC-db organization (file *ForC-EFDB_mapping.csv* available at [https://github.com/forc-db/IPCC-EFDB-integration/blob/main/doc/ForC-EFDB_mapping/ForC-EFDB_mapping.csv](https://github.com/forc-db/IPCC-EFDB-integration/blob/main/doc/ForC-EFDB_mapping/ForC-EFDB_mapping.csv)). 
```{r table_ForCfieldmapping, eval = TRUE, echo=FALSE, warning=FALSE}
library(readr)
library(stringr)
library(flextable)

# Read the data
ForC_variables_included <- read.csv("figures_tables/ForC_fields_included.csv")

# Fix column names
names(ForC_variables_included) <- stringr::str_replace_all(names(ForC_variables_included), "\\.", " ")
column_names <- names(ForC_variables_included) # Define column names
column_names[5] <- "Required*" # Add asterisk for footnote
names(ForC_variables_included) <- column_names# Set column names of the matrix

# Subset to the first 5 columns
ForC_variables_included <- ForC_variables_included[c(1:5)]

# Create flextable object
ft <- flextable(ForC_variables_included)
ft <- autofit(ft)
ft <- width(ft, 1, width = 1)
ft <- width(ft, 2:5, width = 1.5)

# Apply styling
#ft <- ft %>%
#  bold(part = "header") %>%
#  fontsize(size = 8)

ft

```
\* *Required* field indicates whether the field is required by EFDB: yes = value required; (yes) = input required, missing value acceptable if not reported; auto = present within ForC infrasructure, and therefore will always be exported to EFDB ; (no) = not required for EFDB, but required for ForC and therefore will always be exported to EFDB; no = not required, but exported to EFDB when a value is present.

\clearpage

\newpage
**`r table_variables`. Numbers of ForC records and EFDB submissions by variable.**
```{r table_variables, eval = TRUE, echo=FALSE, warning=FALSE}
C_variables_table <- read.csv("figures_tables/C_variables.csv")
names(C_variables_table) <- stringr::str_replace_all(names(C_variables_table), "\\.", " ")
C_variables_table[is.na(C_variables_table)]<- "" #replace NAs with blanks

library(flextable)
set_flextable_defaults(font.family = "Arial", font.size = 8) # set font

ft <- flextable(C_variables_table[,-1])
ft <- autofit(ft)
ft <- width(ft, 1, width = 2)
ft <- width(ft, 2:5, width = 1)
ft <- bold(ft, i = 1, j = NULL, bold = TRUE, part = "header") # put title in bold
ft <- bold(ft, i = which(C_variables_table$bold), j = NULL, bold = TRUE, part = "body") # put categories in bold
ft
```