### 2024-07-03 doc/manuscript/response-to-reviews.Rmd
worked on response to reviews

---
title: ''
output: word_document
bibliography: references.bib #This can be generated by exporting BibTex file from Zotero
csl: new-phytologist.csl

---
```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)

#Tables
table_pools =1
table_recommendations = 2

#Figures
fig_variable_mapping=1
fig_map=2
fig_climate=3
fig_histograms=4

#Appendices
appendix_ForC_updates = "Methods S1"
appendix_IPCC_definitions = "Notes S1"
table_ForCchanges = "Table S1"
table_ForCfieldmapping = "Table S2"
table_variables ="Table S3"
```

Dear Editor:

Below please find our responses to reviews for our manuscript ' Informing forest carbon inventories under the Paris Agreement using ground-based forest monitoring data' (PPP-MT-2024-00903). We have addressed all points, as detailed below.

Sincerely,

Kristina Anderson-Teixeira (on behalf of all coauthors)

\newpage

# Response to Reviews

## Referee: 1

The manuscript presents a process developed for preparing data for submission to the Emission Factor Database (EFDB) and methods to assess the data in the Global Forest Carbon Database (ForC version 4.0) relevant to EFDB. The variables relevant to IPCC and EFDB were identified and listed.
Results describe the databases and analyse some basic elements of the records they contain. Data contribution to ForC is mapped.

The aims of the study are clearly identified and very relevant to the community, and the idea itself of improving scientist’s contribution to international databases appreciable. Thus, there are good reasons to publish.

**Thank you.**

The results section is however disconcerting, partly because it is not clearly structured, and partly because it is quite unclear what point authors wanted to make out of the few graphs presented.

*(NEEDS RESPONSE)*

The first paragraph is about records with deficiencies, a fraction being “deemed not applicable to EFDB…”. This is not informative at all in absence of a series of well-identified examples. Quite verbose but not that useful.

*(NEEDS RESPONSE. We did not receive this information directly. We will have to figure out which records are missing)*

Another example, the map (Fig 2) is fine as an illustration or proof of the global participation to the efforts of reporting, but it doesn’t help identify the gaps in the current reporting. Some vegetation types are well represented, others have a serious deficit. Identifying and quantifying this, for instance in terms of report per area, would have been a more eloquent and useful than a map. The imbalance in the data goes for the below/aboveground compartments too.

**We have made some of the suggested revisions (detailed below) to improve the assessment of data representativeness and consider this to be an important improvement. At the same time, we also wish to emphasize that the goal of this paper was to document the process and provide an overview of the types of data submitted, as opposed to submitting data that adequately represents all forest types and documenting their representativeness. The latter -- to the extent that it would even be possible -- would be a massive project far beyond the scope of this paper. Therefore, while improving the assessment of data representativeness, we believe it would be overkill to analyze all aspects of representativeness.**

**We have added a graph to the map (Fig. `r fig_map`) showing the proportional breakdown by FAO ecozone of (1) area with tree cover, (2) sites with records submitted to EFDB, (3) records submitted to EFDB. We have also added a statement on relative representation by biome: "Relative to forested area, submitted data under-represented the tropics and over-represented temperate and boreal biomes (Fig. `r fig_map`)."**

**To assess the data distribution within in the global climate space, we have added a Whittaker-style diagram showing ForC sites and those with submitted data within the MAT-MAP climate space (new Fig. `r fig_climate`). We have also added a statement on climate representativity: "Submitted data were concentrated in climates that are relatively common over Earth's land area (Fig. `r fig_climate`)." **

**Regarding imbalance among variable types, `r table_variables` provides a complete record of the numbers of records in ForC and submitted to EFDB. We feel that further analyzing by categories such as ecozone or age would be overkill, make the manuscript too long, and of little interest to most readers.**

Fig. 3 *(now 4)* isn’t conclusive either because ecozones have different sizes, so the number of reports on itself is not a sufficient to evaluate the coverage. The representativity of the data within an ecozone depends on the variability between records: if they all cover the same age classes, again, the representativity is bad.

*https://github.com/forc-db/IPCC-EFDB-integration/issues/63*

*Again, we feel that deeper analysis of the representativity would be overkill....*

Consequently, the “Database needs” as discussed L297 are not convincingly demonstrated by the results which are limited to just frequencies. Overall the discussion seems somewhat verbose.

*(NEEDS RESPONSE)*

Some parts of the writing are a bit frustrating, as they suggest that authors have some relevant information that they do not share:

**We'd like to emphasize that all data, code, planning documents, and even our internal discussions are documented in the associated public GitHub repository (https://github.com/forc-db/IPCC-EFDB-integration), as explained in our data availability statement. We have done our best to present relevant information while keeping the manuscript concise and as interesting as possible. We appreciate the reviewer having pointed out places where more information is desirable, and have modified the text to include it.**

L156 “We identified ForC variables that were relevant to the IPCC…”
Great, but how? The reasoning behind the identification is certainly of interest to scientists or practitioners.

**We have modified this statement to read, "Working in consultation with the EFDB Technical Support Unit and referencing IPCC guidance [@ipcc_good_2003; @ipcc_2006_2006; @ipcc_2019_2019], we identified ForC variables that were relevant to the IPCC methodology and EFDB (Fig. `r fig_variable_mapping`, `r appendix_IPCC_definitions`)."**

L226 “To submit complete, reviewed ForC records into EFDB, we created R scripts to restructure ForC records and populate EFDB’s bulk import form.”
What are those scripts?
Why not share these scripts?

**These scripts are publicly available in the associated GitHub repository (https://github.com/forc-db/IPCC-EFDB-integration) and also archived in Zenodo. This is explained in our data availability statement, but that fact is easy to miss. To ensure that this is clear to readers, we have modified this statement to read: "To submit complete, reviewed ForC records into EFDB, we created R scripts to restructure ForC records and populate EFDB's bulk import form (publicly available; see Data Availability Statement).**

L300. So far, only ~7% of the EFDB-relevant data in ForC have been submitted to EFDB.
How do you get that number? What are those 7% and what consequences does this have?
Providing concrete examples of mistakes to avoid would have been very appreciable, and made the text sound less theoretical.

*(Provide a bit more detail, addressing questions.)*

The discussion section “Data reporting needs” should be named “Recommendations of good practices for reporting C stocks and stock changes”, or something more in line with the content itself.

*(NEEDS RESPONSE)*

The summary lists up a point on human’s influence on earth’s climate, which is also heavily elaborated in the introduction. The emphasis on human’s influence on forests and climate are disproportionated and distracting (i.e. not appropriate here), since the manuscript is not dedicated to studying it as it solely focuses on scientists reporting C stocks evaluations. A simple statement that reporting C stocks is an obligation but still poses difficulties would be more honest and efficient. Removing the first paragraph of the introduction, which is totally disconnected, would allow to shorten a bit the overall too long introduction.

*(NEEDS RESPONSE)*

L239: “The 37% attenuation between records we reviewed and those submitted…”
Why “attenuation”?, it seems to be a mere discrepancy, or deficit.

**We have changed this word to "discrepancy".**

In conclusion, although the aims and principle seem simple, the results are both vague and expeditive, resulting in a rather unconvincing article. Some efforts to make results more relevant, and some more details on key aspects would help.

*(NEEDS RESPONSE)*

# Referee: 2

This manuscript is a timely and relevant paper about developing a standardized protocol to input forest carbon data into the IPCC emissions database. The authors identified two main problems: the lack of authors submitting their published forest carbon data into the IPCC Emission Factor Database, and some papers not fully reporting all information (pools, fluxes, methodology) necessary to submit in the EFDB). This team has developed a robust and standardized methodology for first aggregating forest carbon data into their already developed (but modified for this new methodology) ForC database, followed by inputting ForC data into the IPCC database. The lack of available, standardized forest carbon data is a real problem as the need to monitor ecosystem carbon stocks and fluxes becomes more crucial in the face of climate change– the methodology developed in this paper has high societal impact to better fill in data gaps in a globally important IPCC database.

**Thank you.**

This paper is high quality and well written with reliable findings. The authors used appropriate techniques for both the ForC database and the methodology for semi-automated data submission into the EFDB. Figure 1 is easy to follow and shows how detailed and robust the ForC to EFDB framework is.

**Thank you.**

I appreciate the authors recommendations in the discussion section (they are all relevant, warranted, and important) but I do think some of these recommendations could be more strongly worded. The authors have described large challenges with populating the EFDB, and have developed a robust solution (and identified other, easier solutions). I would like to see the authors clear and strong recommendations in this section. Some of the suggestions would require substantial increases in funding and infrastructure (line 328-333), which are huge challenges. I think that section has been well discussed in the discussion. However, the next paragraph discusses the lack of variables available because they have either not been calculated or just not presented in papers (although they may have been available to the authors) (lines 334-346). I think recommendations in this section could be stronger and more specific – it is very easy for authors to provide data they have already collected/calculated, or to complete relatively simple allometric calculations. Are there ways this could be required or incentivized (this question may be out of scope for this paper)? A more specific recommendation in the next paragraph (lines 348-362) would also be warranted.

*(NEEDS RESPONSE)*

\clearpage
# References {-}